### 2019-02-25 paper/intro-notes.tex
Add paragraph on son preference and spacing research

Birth spacing has long served as a measure of son preference, with strong
son preference typically associated with shorter birth intervals the
fewer sons a family has \citep{ben-porath76b,Leung1988}.
However, the introduction of prenatal sex determination has the potential to 
change the relationship between son preference and birth spacing fundamentally.
Some of the couples who before had shorter birth spacing because of son preference,
might instead resort to sex-selective abortions, and each abortion naturally 
increases the interval between births.
As a result, we may now observe \emph{longer} average spacing for families with strong son 
preference, precisely because it leads them to use sex selection.
To further complicate matters, families who---for one reason or another---do 
not use prenatal sex selection but still have a preference for son may continue to have 
shorter birth spacing than those with a lower preference for sons.

@@

There is amble evidence that son preference has affected individual families' fertility
decisions in India and elsewhere.
It is, for example, clear that a significant proportion of Indian families follow a
son preference based differential stopping behavior, where, for a given number of
children, they are more likely to have an additional child if they have not yet
reached their preferred number of sons 
\citep{repetto72,Das1987,Arnold1997,clark00,Basu2010,Barcellos2014}.
A smilar pattern is found in many other countries 
\citep[see, for example][]{larsen98,filmer09,Altindag2016}.%
\footnote{
One potential effect of this behavior is that girls tend to end up in larger families, 
which may have negative impacts on survival and human capital investments, even in 
the absence of direct discrimination \citet{Jensen2003,Kugler2017}.
}

Of more direct interest here is differential \emph{spacing} behavior, both because 
it has been used as a measure for son preference and because birth spacing may directly
impact health and other outcomes for both children and mother.

My proposed method allows for the time since the previous 
birth to affect the decision on sex selection.
By examining under what circumstances sex selection decisions change with 
spacing, we can draw a more nuanced picture of the degree of son preference.

[son preference and spacing - evidence]

@@

[falling fertility and introduction of sex selection]

[Predicted changes in spacing]

@@

>>>>

Birth spacing has long served as a measure of son preference, with strong
son preference typically associated with shorter birth intervals the
fewer sons a family has \citep{ben-porath76b,Leung1988}.
However, the introduction of prenatal sex determination potentially  
changed the relationship between son preference and birth spacing fundamentally.
The stronger the son preference, the more likely couples are to resort to sex-selective 
abortions, and each abortion naturally increases the interval between births.
As a result, we may now observe \emph{longer} average spacing for families with stronger 
son preference, precisely because it makes them more likely to use sex selection.
To further complicate matters, families with a preference for son who---for one reason or 
another---do not use prenatal sex selection may continue to have shorter birth spacing 
than those with a lower preference for sons.

@@

There is amble evidence that son preference has affected fertility decisions in India and 
elsewhere.
It is, for example, clear that a significant proportion of Indian families follow a
son preference based differential stopping behavior, where, for a given number of
children, they are more likely to have an additional child if they have not yet
reached their preferred number of sons 
\citep{repetto72,Das1987,Arnold1997,arnold98,clark00,Basu2010,Barcellos2014}.
A smilar pattern is found in many other countries 
\citep[see, for example][]{larsen98,filmer09,Altindag2016}.%
\footnote{
One potential effect of this behavior is that girls tend to end up in larger families, 
which may have negative impacts on survival and human capital investments, even in 
the absence of direct discrimination \citet{Jensen2003,Kugler2017}.
}

[emphasize that birth spacing is a superset of DSB?]
Of more direct interest here is, however, the potential for differential \emph{spacing} 
behavior, 
both because it has been used as a measure for son preference and because birth spacing 
may directly impact health and other outcomes for both children and mother.
In India the sex of the last-born child significantly impact the duration to next birth,
with the expected birth interval about three percent longer if the last-born was son than
if it was a daughter \citep{Bhalotra2008,Kumar2016}.
Similarly, ethic Indians in South Africa also show a longer duration after the birth of a 
son than after a daughter \citep{Gangadharan2003}.%
\footnote{
Ethic Indians in Malaysia do, however, show little evidence of son preference, although
the sample sizes are very small \citep{Pong1994}.
}
India is not alone in showing an effect of sex composition on subsequent birth spacing.
In both Bangladesh and Vietnam the more boys a family has the longer the expected duration 
to the next birth \citep{Haughton1995,Haughton1996,Rahman1993,Soest2018}.
Outside Asia the evidence is more mixed.
North Africa shows shorter spacing in the absense of sons, while a similar pattern does
not exists in Sub-Saharan Africa \citep{Rossi2015}.%
\footnote{
A possible exception is Senegal, where women with a higher risk of widowhood show
shorter birth spacing until they have secured a son \citep{Lambert2016}.
}

[how is spacing achieved?]
This longer spacing 
\citep{Rahman1993}
\citep{Jayachandran2011}

[Education and socio-economic status and spacing]

@@

@@

[Predicted changes in spacing with sex selection]
